---
title: "Parasitemia and Incidence"
author: "jp"
date: "11/3/2016"
output:
  html_document: default
  word_document: default
---


```{r, include = FALSE, cache=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache = FALSE)

library(data.table)
library(tidyverse)
library(stringr)
library(scales)
library(forcats)
library(survey)
library(gridExtra)
library(countrycode)
library(readr)
library(readxl)
library(DT)
library(plotly)
library(zoo)

```


# A Comparison of malaria parasitemia (from household surveys) with incidence data (from WHO, World Malaria Report)


## Survey data 

```{r,  include = FALSE}

source("file_list.r")

subsahara = countrycode_data %>%
  filter( region %in% c("Middle Africa", "Western Africa", "Eastern Africa", "Southern Africa")) %>% 
  select( country.name, iso3c, iso2c ) %>%
  filter( !country.name %in% "Lesotho") 


files = files %>% 
  mutate( country.name = countrycode(country, "country.name", "country.name")) %>%
  filter( survey!="", year >= 2006 , country.name %in% subsahara$country.name)

# not_needed = c('Bangladesh', 'Pakistan', 'Philippines', 'Comoros')

survey_list = files %>% count(country, survey, year) %>% select(-n) %>% arrange(desc(year))

nsurveys = nrow(survey_list)
ncountries = length( unique( files$country ) )

datatable(survey_list)

```

```{r, include = FALSE}
# select variables for this analysis, v
variables =  c('v000', 
       'v001', 'v002', 'v003', 'v005', 'v021', 'v023', 'v024', 'v025', 
       'weight.c', 'weight.hm', 'weight.w', 'weight.h', 
       'hv001',  'hv002', 'hv003',  'hv005', 'hv007', 'hv021', 'hv022', 'hv023', 'hv024', 'hv025',
       'hvidx', 'b16',
       'h22', 'hv105', 'h47', 
       'hml16', # age in years
       'hml32', 'hml35' )

```

```{r, include = FALSE, eval= FALSE}
source("dhs_load_survey.R")

nsurveys = nrow(survey_list)

survey_list = survey_list %>% arrange( desc(year) )
index = row.names(survey_list)

p <- progress_estimated(nsurveys)

x <- vector("list", length( nsurveys)) 

translate = FALSE  # optional translation of values to label.  
  # NB: Does not work well for all variables

for (i in seq_along( index ) ){
  
   p$pause(0.1)$tick()$print()
  
  .country = survey_list[i, ]$country
  .iso3c = countrycode( survey_list[i, ]$country, "country.name", "iso3c")
  .survey = survey_list[i, ]$survey
  .year = survey_list[i, ]$year
  .survey_name = paste(.survey, .year)

  print( paste(.country, .survey, .year))
  
  svy <- try( silent = TRUE,
    load_survey_object( design = FALSE,
                        # printout = TRUE ,  # un comment when testing
                        vars = variables, 
                        .country = .country, 
                        .survey = .survey,
                        .year = .year)
  )
  
  #TODO FILTER RECORDS--apply filter, like keeping only children with parasitemia, is not unnecessarily large

  if ( class(svy) == "try-error" ){ 
    next()
    }

  # returns: svy.h, svy.c, svy.w, svy.hm, vars_x, x
  # svy.h <- svy[[1]]
  # svy.c <- svy[[2]]
  # svy.w <- svy[[3]]
  # svy.hm <- svy[[4]]
  has_vars = svy[[5]]
  x[[i]] = svy[[6]] 
  
  x[[i]]$country = rep( .country, nrow(x[[i]]) )
  x[[i]]$iso3c = rep( .iso3c, nrow(x[[i]]) )
  x[[i]]$year = rep( .year, nrow(x[[i]]) )
  x[[i]]$survey = rep( .survey, nrow(x[[i]]) )

  # TODO: optional- relate dictionary to survey                    )

  # # TODO: only for columns with >4 values (others easier to keep as numeric)
  # # Deriving this table should happen once, before loop
  # v = dd %>% filter( tolower(Item_Name) %in% variables) %>% group_by( Item_Name, label ) %>% 
  #   summarise( n = n()) %>% group_by( Item_Name) %>% count(Item_Name) %>% filter(nn>4) 

  if (translate){
    
      # refine dictionary for this country-survey 
      dict = dd %>% filter_( ~country == .country , ~survey == .survey, ~year == .year,
                        ~Item_Name %in% toupper(names(has_vars))
      )
     
    y = x[[i]]
  
    for (col in 1:length(colnames(y))){
      column = tolower( colnames(y)[col] )
      
      dict_var = dict %>% 
        mutate( item_name = tolower(Item_Name)) %>%
        filter_( ~item_name == column ) %>% # , ~file == "Household Member Recode"
        count( value, label) %>% select(-n)
      
      old_value = y[, col]
      new_value = dict_var[ match( old_value, dict_var$value ), ]$label
      y[, col] = ifelse( is.na( new_value), old_value, new_value)
    }
      x[[i]] = y
}

  # if (is.null(X) ){ X <- x; next()}
  # X = rbindlist( list(X, x), use.names = TRUE, fill = TRUE)
  
}
  
# X = dplyr:: bind_rows( x )  # will throw error is column types dont match 
X = rbindlist(x, fill = TRUE) %>% rownames_to_column("id")

saveRDS(X, "rdt_microscopy.rds")

# object.size(X)

```

```{r}
x = readRDS("rdt_microscopy.rds")
```

## subset to those with microscopy OR RDT results: 

```{r, include= FALSE, message=""}

dataset = x %>% 
  mutate( startyear = as.integer( substr( year, 1, 4 ) ),
            endyear = as.integer(
              ifelse( nchar(year)>4, 
                              paste0("20", substr( year, 6, 7 )), 
                              substr( year, 1, 4 ))
            )
            ) %>%
  filter( 
          # endyear >= 2006, 
          # iso3c %in% subsahara$iso3c ,
          !is.na(weight.c) ,  # only those with weight 
          ( !is.na(hml32) | !is.na(hml35) ) # non-missing RDT or slide
          ) %>%
  rename( 
            psu = hv021,
            region = hv024,
            strata = hv023 ,
            slide = hml32, 
            rdt = hml35 
            ) %>%
  as.data.table()

dataset.list = dataset %>% count( country, survey, year ) %>% arrange(desc(year) )

cat("There are", nrow(dataset.list), "surveys of", comma(nrow(dataset)), "individuals having RDT or microscopy results")

datatable( dataset.list %>% select(-n) )

```


## Average Test Positivity by Country

```{r}

  d_strata = 
  dataset %>% 
  # summarise by region, psu
    group_by( country, survey, year, startyear, endyear, 
              region, psu ) %>% # hv024 = region, hv023 = sample strata
    summarise( n = n(),
               n_slide = sum( slide %in% 0:1 ),
               n_rdt = sum( rdt %in% 0:1 ),
               slide_pos  = sum(slide %in% 1),
               rdt_pos = sum(rdt %in% 1),
               slidepos_rdtneg = sum( ifelse( slide %in% 1 & rdt %in% 0, 1, 0) ),
               slideneg_rdtpos = sum( ifelse( slide %in% 0 & rdt %in% 1, 1, 0) ),
               discordant = sum(slidepos_rdtneg , slideneg_rdtpos, na.rm = TRUE)
               ) %>%
    mutate( `%slide` = ifelse( n_slide == 0, NA, slide_pos / n_slide )  ,
            `%rdt` = ifelse( is.na(n_rdt), NA, rdt_pos / n_rdt ),
            `%slidepos_rdtneg` = ifelse( is.na(n), NA, slidepos_rdtneg / n) ,
            `%slideneg_rdtpos` = ifelse( is.na(n), NA, slideneg_rdtpos / n) ,
            `%discordant` = ifelse( is.na(n), NA, discordant / n ),
            id = row_number(),
            iso2c = countrycode( country, "country.name", "iso2c")
            ) 

d_country = d_strata %>%
  group_by(country, iso2c, survey, year, startyear, endyear) %>%
  summarise(
    n = sum(n),
    n_rdt = sum(n_rdt),
    n_slide = sum(n_slide),
    rdt = 100 * sum(rdt_pos) / n_rdt ,
    slide = 100* sum(slide_pos) / n_slide ) %>%
  mutate(
    id = row_number()
    
  ) %>%
  as.data.frame() 


```

```{r, include=FALSE}

    gg = d_country %>% 
      gather_( "Test", "pct", gather_cols = c('rdt', 'slide') ) %>%
      ggplot( aes( factor(endyear), pct, label = country, color = country, shape = Test) ) +
      geom_point( ) +
      guides( color = FALSE, shape = FALSE ) +
      labs( x = "Year", y = "% Test Positive", title = "Mean positivity rate of RDT and microscopy") 


  ggplotly( gg )

```

##  most recent survey

```{r}

last_parasitemia = 
  d_country %>%
  filter( rdt > 0 | slide > 0 ) %>%
  group_by( country ) %>%
  arrange( -endyear ) %>%
  filter( row_number() == 1 ) %>%
  ungroup %>% arrange( country ) %>%
  rowwise() %>%
  mutate(
    prevalence = ifelse( sum(rdt, slide, na.rm = TRUE) == 0, NA, 
                         sum(rdt, slide, na.rm = TRUE) /2 )
  )

```

```{r}
    gg = last_parasitemia %>% 
      ggplot( aes( factor(endyear), prevalence, label = country, color = country) ) +
      geom_point( ) +
      guides( color = FALSE, shape = FALSE ) +
      labs( x = "Year", y = "% Test Positive", title = "Mean positivity rate of RDT and microscopy") 


  ggplotly( gg )

```


# Number of cases reported to WHO

```{r, include = FALSE}


# wmr_reports <- read_excel("../WMR/SEE_WMR2016_MalariaData Oct 6.xlsm", 
#     sheet = "Cases and Deaths") 

wmr_reports <- read_excel("C:/Users/bzp3/dropbox/_Malaria/Projects/WMR/SEE_WMR2016_MalariaData_ML.xlsx", 
    sheet = "Under5")

# View( wmr_reports %>% filter( Country == "Sierra Leone") )

wmr_cases = wmr_reports %>%
  rename( rpt_presumed_confirmed= `malaria cases (presumed + conf) total`,
          calc_presumed_confirmed = `calc - indigenous presumed and confirmed`,
          calc_confirmed = `calc - total positive (micr+rdt)`,
          endyear = Year, country = Country, iso2c = ISO_Ctry) %>%
  filter( iso2c %in% subsahara$iso2c , endyear >= 2006  ) %>%
  select( country, iso2c, endyear, 
          rpt_presumed_confirmed, calc_presumed_confirmed, calc_confirmed 
          ) %>%
  mutate( 
    presumed_confirmed = ifelse( is.na(rpt_presumed_confirmed), 
                                 calc_presumed_confirmed, 
                                 rpt_presumed_confirmed) 
    )

# If missing, fill in from previous year

# wmr_cases %>% kable

wmr_cases. = wmr_cases %>%
  group_by( country, endyear) %>% 
  na.locf %>% ungroup %>%
  mutate( endyear = as.integer(endyear),
          presumed_confirmed = as.integer(presumed_confirmed),
          calc_confirmed = as.integer( calc_confirmed ),
          percent_confirmed = calc_confirmed / presumed_confirmed
          )

wmr_pop <- read_excel("../WMR/SEE_WMR2016_MalariaData Oct 6.xlsm", 
    sheet = "Population")[1:10] %>%
  select( Country, ISO_Ctry, Year, Total, `Population risk rate (Total)`) %>%
  filter( ISO_Ctry %in% subsahara$iso2c , Year >= 2006  ) %>%
  rename( pop = Total, endyear = Year, country = Country, iso2c = ISO_Ctry) %>%
  mutate(
    pop_risk = pop * `Population risk rate (Total)`
  )

wmr_incidence = inner_join(wmr_cases., wmr_pop) %>%
  mutate( inc_total = presumed_confirmed  / pop_risk ,
          inc_confirmed = calc_confirmed  / pop_risk )


```

```{r}
wmr_incidence %>% 
  # filter(  endyear == 2015 ) %>%
  mutate(
    presumed_confirmed = comma( presumed_confirmed ),
    calc_confirmed = comma( calc_confirmed),
    pop_risk = comma( pop_risk , digits = 0)
    ) %>%
  arrange( -inc_total ) %>% 
  select( -pop,  -`Population risk rate (Total)`, 
         -calc_presumed_confirmed, -rpt_presumed_confirmed) %>%
  datatable( caption = "Total malaria cases reported to WHO")

```

## Note that cases are reported largely from the public reporting system (e.g.RHIS).  the proportion of persons going to a public clinic (versus private) varies between countries.  To better compare the incidence between countries, it may be reasobnable to adjust the rates by dividing by the percent of patients that go to public clinics.  Surveys have the percent of children who access care by provider type, which could be a proxy for percent of all patients.  (World bank household surveys may have better data for adults) . 

# Incidence versus Prevalence (most recent survey)

```{r, include = FALSE}

d = left_join( wmr_incidence , last_parasitemia %>% select(-country), by = c("iso2c", "endyear") ) %>%
  filter( !is.na(id) ) %>%
  mutate( label = paste( country, endyear ) )

```


```{r}

gg = ggplot(d, aes( prevalence, inc_total, color = country , label = label)) +
  geom_point() +
  labs( x = "prevalence (most recent survey)", y = "Presumed and confirmed cases (1/PYO)" ,
        title = "Incidence v. Prevalence \n(hover to see details)") +
  guides( color = FALSE )

ggplotly( gg , tooltip = "label")

```

# Incidence versus Prevalence (All surveys)

```{r, include = FALSE}

all_parasitemia = 
  d_country %>%
  filter( rdt > 0 | slide > 0 ) %>%
  group_by( country ) %>%
  rowwise() %>%
  mutate(
    prevalence = ifelse( sum(rdt, slide, na.rm = TRUE) == 0, NA, 
                         sum(rdt, slide, na.rm = TRUE) /2 )
  )

d_all = left_join( wmr_incidence , all_parasitemia ) %>%
  filter( !is.na(id) ) %>%
  mutate( label = paste( country, endyear ) )

# View( d_all )

```


```{r}

gg = ggplot(d_all, aes( prevalence, inc_total, color = country , label = label)) +
  geom_line( aes( group = country )) +
  geom_point( aes( size = endyear )) +
  labs( x = "prevalence (most recent survey)", y = "Presumed and confirmed cases (1/PYO)" ,
        title = "Incidence v. Prevalence \n(hover to see details)") +
  guides( color = FALSE )

ggplotly( gg , tooltip = "label")

```

# Incidence--confirmed cases only--versus Prevalence  (most recent survey)

```{r}

gg = ggplot(d, aes( prevalence, inc_confirmed, color = country, label = label )) +
  geom_point() +
  labs( x = "prevalence (most recent survey)", 
        y = "Confirmed cases (1/PYO)" ,
        title = "Incidence-confirmed cases v. Prevalence \n(hover to see details)") +
  guides( color = FALSE )

ggplotly( gg , tooltip = "label")

```

# Incidence--confirmed cases only--versus Prevalence  (All surveys)

```{r}

gg = ggplot(d_all, aes( prevalence, inc_confirmed, color = country, label = label )) +
  geom_line( aes( group = country )) +
  geom_point( aes( size = endyear )) +
  labs( x = "prevalence (most recent survey)", 
        y = "Confirmed cases (1/PYO)" ,
        title = "Incidence-confirmed cases v. Prevalence \n(hover to see details)") +
  guides( color = FALSE )

ggplotly( gg , tooltip = "label")

```

# compare number of cases reported to WHO and number of cases reported in MOP

```{r, include=FALSE}

load("rsi2.rda")

rsi2 = rsi2 %>% mutate( iso2c = countrycode(country, "country.name", "iso2c"))

display_cols = c("country", "Cases: total", "Cases: confirmed", "Cases: suspect", "Completeness", "Deaths: Total", "TPR")

rsi2[, display_cols]

# case count plot ####


gg = ggplot(rsi2, aes(x = cases, y = fct_reorder(country, cases))) + 
   geom_point(stat = 'identity') +
   scale_x_continuous(labels = comma) +
   scale_y_discrete("Country")

ggplotly( gg )

```

```{r}

mop_v_wmr = inner_join( rsi2, wmr_incidence %>% filter( endyear == 2015 ), by = "iso2c")

gg = ggplot(mop_v_wmr, aes( cases, as.numeric(calc_presumed_confirmed), label = country.x ) ) +
  geom_point() +
  labs( x = "MOP", y = "WHO") + 
  geom_abline(intercept = 0, slope = 1, alpha = .25)
  

ggplotly( gg )

```

