---
title: "Slide vs RDT"
runtime: shiny
output: 
  # ioslides_presentation 
  html_document:
    css: shinyprezcss.css
    fig_height: 6
    fig_width: 10
    number_sections: yes
    toc: yes
---

```{r, include = FALSE, cache=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, cache = FALSE)

library(data.table)
library(tidyverse)
library(stringr)
library(scales)
library(survey)
library(gridExtra)
library(countrycode)
library(readr)
library(readxl)
library(ggvis)  

```


# A Comparison of DHS and MIS malaria parasitemia estimates by microscopy (slides) with RDT 


## Main data 

```{r,  include = FALSE}

source("file_list.r")

subsahara = countrycode_data %>%
  filter( region %in% c("Middle Africa", "Western Africa", "Eastern Africa", "Southern Africa")) %>% 
  select( country.name, iso3c, iso2c ) %>%
  filter( !country.name %in% "Lesotho") 


files = files %>% 
  mutate( country.name = countrycode(country, "country.name", "country.name")) %>%
  filter( survey!="", year >= 2006 , country.name %in% subsahara$country.name)

# not_needed = c('Bangladesh', 'Pakistan', 'Philippines', 'Comoros')

survey_list = files %>% count(country, survey, year) %>% select(-n) %>% arrange(desc(year))

nsurveys = nrow(survey_list)
ncountries = length( unique( files$country ) )

kable(survey_list, row.names = TRUE)

```

```{r, include = FALSE}
# select variables for this analysis, v
variables =  c('v000', 
       'v001', 'v002', 'v003', 'v005', 'v021', 'v023', 'v024', 'v025', 
       'weight.c', 'weight.hm', 'weight.w', 'weight.h', 
       'hv001',  'hv002', 'hv003',  'hv005', 'hv007', 'hv021', 'hv022', 'hv023', 'hv024', 'hv025',
       'hvidx', 'b16',
       'h22', 'hv105', 'h47', 
       'hml16', # age in years
       'hml32', 'hml35' )

```

```{r, include = FALSE, eval= FALSE}
source("dhs_load_survey.R")

nsurveys = nrow(survey_list)

survey_list = survey_list %>% arrange( desc(year) )
index = row.names(survey_list)

p <- progress_estimated(nsurveys)

x <- vector("list", length( nsurveys)) 

translate = FALSE  # optional translation of values to label.  
  # NB: Does not work well for all variables

for (i in seq_along( index ) ){
  
   p$pause(0.1)$tick()$print()
  
  .country = survey_list[i, ]$country
  .iso3c = countrycode( survey_list[i, ]$country, "country.name", "iso3c")
  .survey = survey_list[i, ]$survey
  .year = survey_list[i, ]$year
  .survey_name = paste(.survey, .year)

  print( paste(.country, .survey, .year))
  
  svy <- try( silent = TRUE,
    load_survey_object( design = FALSE,
                        # printout = TRUE ,  # un comment when testing
                        vars = variables, 
                        .country = .country, 
                        .survey = .survey,
                        .year = .year)
  )
  
  #TODO FILTER RECORDS--apply filter, like keeping only children with parasitemia, is not unnecessarily large

  if ( class(svy) == "try-error" ){ 
    next()
    }

  # returns: svy.h, svy.c, svy.w, svy.hm, vars_x, x
  # svy.h <- svy[[1]]
  # svy.c <- svy[[2]]
  # svy.w <- svy[[3]]
  # svy.hm <- svy[[4]]
  has_vars = svy[[5]]
  x[[i]] = svy[[6]] 
  
  x[[i]]$country = rep( .country, nrow(x[[i]]) )
  x[[i]]$iso3c = rep( .iso3c, nrow(x[[i]]) )
  x[[i]]$year = rep( .year, nrow(x[[i]]) )
  x[[i]]$survey = rep( .survey, nrow(x[[i]]) )

  # TODO: optional- relate dictionary to survey                    )

  # # TODO: only for columns with >4 values (others easier to keep as numeric)
  # # Deriving this table should happen once, before loop
  # v = dd %>% filter( tolower(Item_Name) %in% variables) %>% group_by( Item_Name, label ) %>% 
  #   summarise( n = n()) %>% group_by( Item_Name) %>% count(Item_Name) %>% filter(nn>4) 

  if (translate){
    
      # refine dictionary for this country-survey 
      dict = dd %>% filter_( ~country == .country , ~survey == .survey, ~year == .year,
                        ~Item_Name %in% toupper(names(has_vars))
      )
     
    y = x[[i]]
  
    for (col in 1:length(colnames(y))){
      column = tolower( colnames(y)[col] )
      
      dict_var = dict %>% 
        mutate( item_name = tolower(Item_Name)) %>%
        filter_( ~item_name == column ) %>% # , ~file == "Household Member Recode"
        count( value, label) %>% select(-n)
      
      old_value = y[, col]
      new_value = dict_var[ match( old_value, dict_var$value ), ]$label
      y[, col] = ifelse( is.na( new_value), old_value, new_value)
    }
      x[[i]] = y
}

  # if (is.null(X) ){ X <- x; next()}
  # X = rbindlist( list(X, x), use.names = TRUE, fill = TRUE)
  
}
  
# X = dplyr:: bind_rows( x )  # will throw error is column types dont match 
X = rbindlist(x, fill = TRUE) %>% rownames_to_column("id")

saveRDS(X, "rdt_microscopy.rds")

# object.size(X)

```

```{r}
x = readRDS("rdt_microscopy.rds")
```

## subset to available surveys with microscopy and RDT results: 

```{r}

dataset = x %>% 
  mutate( startyear = as.integer( substr( year, 1, 4 ) ),
            endyear = as.integer(
              ifelse( nchar(year)>4, 
                              paste0("20", substr( year, 6, 7 )), 
                              substr( year, 1, 4 ))
            )
            ) %>%
  filter( 
          # endyear >= 2006, 
          # iso3c %in% subsahara$iso3c ,
          !is.na(weight.c) ,  # only those with weight 
          !is.na(hml32) , 
          !is.na(hml35)
          ) %>%
  rename( 
            psu = hv021,
            region = hv024,
            strata = hv023 ,
            slide = hml32, 
            rdt = hml35 
            ) %>%
  as.data.table()

dataset.list = dataset %>% count( country, survey, year ) %>% arrange(desc(year) )

cat("There are", nrow(dataset.list), "surveys comprising", comma(nrow(dataset)), "individuals having both slide and RDT result.")

kable( dataset.list %>% select(-n) , row.names = TRUE)

```

## data dictionary

```{r, message=FALSE, warning=FALSE, include = FALSE}
# get dictionary vales from children's file

  # dd <-readRDS("dictionaryNew.rds")  # about 6 sec

#Which files have treatment variables h37 and ml13?  both in children's file
# count( filter(dd, grepl( "H37|ML13", Item_Name )), file)

  dd = readRDS("dictionaryNew.rds")  %>% ungroup %>%
    filter( Item_Name %in% toupper(variables),
            country %in% x$country,
            survey %in% x$survey,
            year %in% x$year
            ) %>%
    mutate(
      startyear = as.integer( substr( year, 1, 4 ) ) ,
      endyear = as.integer(
              ifelse( nchar(year)>4, 
                              paste0("20", substr( year, 6, 7 )), 
                              substr( year, 1, 4 ))
            ),
      Item_Label = tolower( Item_Label ) ,
      Item_Name = tolower( Item_Name)
    ) %>%
    filter( endyear >= 2006 ) 
  
  # View(dd)

```

## Average Test Positivity by Country

```{r}

  # percent forumula that respects zero
  my.percent<- function(x) {
    if(length(x)==1) if(x==0) return(paste0(0,"%") )
    return(percent(x) )
  }

  d_strata = 
  dataset %>% 
  # summarise by region, psu
    group_by( country, survey, year, region, psu ) %>% # hv024 = region, hv023 = sample strata
    summarise( n = n(),
               n_slide = sum( slide %in% 0:1 ),
               n_rdt = sum( rdt %in% 0:1 ),
               slide_pos  = sum(slide %in% 1),
               rdt_pos = sum(rdt %in% 1),
               slidepos_rdtneg = sum( ifelse( slide %in% 1 & rdt %in% 0, 1, 0) ),
               slideneg_rdtpos = sum( ifelse( slide %in% 0 & rdt %in% 1, 1, 0) ),
               discordant = sum(slidepos_rdtneg , slideneg_rdtpos, na.rm = TRUE)
               ) %>%
    mutate( `%slide` = my.percent(slide_pos / n_slide),
            `%rdt` = my.percent( rdt_pos / n_rdt),
            `%slidepos_rdtneg` = my.percent( slidepos_rdtneg / n),
            `%slideneg_rdtpos` = my.percent( slideneg_rdtpos / n),
            `%discordant` = my.percent( discordant / n ),
            id = row_number()
            ) 

d_country = d_strata %>%
  group_by(country, survey, year) %>%
  summarise(
    n = sum(n),
    n_rdt = sum(n_rdt),
    n_slide = sum(n_slide),
    rdt = 100 * sum(rdt_pos) / n_rdt ,
    slide = 100* sum(slide_pos) / n_slide ) %>%
  mutate(
    id = row_number()
  ) %>%
  as.data.frame() 


```

```{r, cache = FALSE}

  survey_label <- function(x) {
      if(is.null(x)) return(NULL)
      paste( d_country[x$id, "country"], d_country[x$id, "survey_name"])
    }
#     
    reference_line = data.frame( x= c(0,85), y = c(0,85))
    
    d_country %>%
      ggvis(~slide, ~rdt) %>%
      layer_model_predictions( model = "lm", se = TRUE) %>%
      layer_points( fill = ~country, key := ~id) %>%
      add_axis("x", title = "Slide % Positive") %>%
      add_axis("y", title = "RDT % Positive") %>%
      # scale_numeric("x", domain = c(0, 85)) %>%
      # scale_numeric("y", domain = c(0, 85)) %>%
      layer_paths( data = reference_line, x = ~x, y= ~y, stroke:="green") %>%
      add_tooltip( survey_label(), "click") 
#
 
# lm regression
    linear = lm( rdt ~ slide, data = d_country)
    # summary(linear)
    `Linear Regression` = as.data.frame(cbind(round(linear$coefficients, 2),
                                              confint(linear))
                                        )
    names(`Linear Regression`) = c("Estimate", "LL", "UL")
    rownames(`Linear Regression`) = c("Intercept", "Slope")

    renderTable( `Linear Regression` )
```

## Surveys Available through DHS^[http://dhsprogram.com/data/available-datasets.cfm]

```{r, echo=FALSE, message= FALSE, cache = FALSE}

  observe({
    if (!is.null(input$myTable) && input$myTable != -1)
      updateTextInput(session, "myTextArea", value = retrieveAbstract(input$myTable))
  })
    
renderTable({ d_country  })

```

## country-region

# compare results of RDT and sLide

```{r}
  
  # summarise by psu
  compare = 
    dataset %>% 
    mutate(
      survey_name = paste(country, survey, year, sep = "_")
      ) %>%
    group_by( country, survey_name, region, psu ) %>%
    summarise( n = n(),
               n_slide = sum( slide %in% 0:1 ),
               n_rdt = sum( rdt %in% 0:1 ),
               slide_pos  = sum(slide %in% 1),
               rdt_pos = sum(rdt %in% 1),
               slidepos_rdtneg = sum( ifelse( slide %in% 1 & rdt %in% 0, 1, 0) ),
               slideneg_rdtpos = sum( ifelse( slide %in% 0 & rdt %in% 1, 1, 0) ),
               discordant = sum(slidepos_rdtneg , slideneg_rdtpos, na.rm = TRUE)
               ) %>%
    mutate( `%slide` = my.percent(slide_pos / n_slide),
            `%rdt` = my.percent( rdt_pos / n_rdt),
            `%slidepos_rdtneg` = my.percent( slidepos_rdtneg / n),
            `%slideneg_rdtpos` = my.percent( slideneg_rdtpos / n),
            `%discordant` = my.percent( discordant / n )
            ) 


```

```{r, echo=FALSE, message=FALSE, cache= FALSE}
    
  survey_label_cr <- function(x) {
      dd = d_cr()
      if(is.null(x)) return(NULL)
      paste( dd[x$id, "country"], 
             "region", dd[x$id, "region"], 
             dd[x$id, "survey_name"], 
             "n=", dd[x$id, "n"])
    }

  d_cr =  
    reactive({
      compare %>%
      group_by(country, survey_name, region) %>%
      summarise(
        n = sum(n),
        n_rdt = sum(n_rdt),
        n_slide = sum(n_slide),
        rdt = 100 * sum(rdt_pos) / n_rdt ,
        slide = 100* sum(slide_pos) / n_slide 
                ) %>%
      as.data.frame() %>%
      mutate( id = row_number()) 
  })    


    d_cr %>% 
      ggvis(~slide, ~rdt ) %>%
      layer_model_predictions( model = "lm", se = TRUE
                               ) %>%
      ungroup() %>%
      layer_points( fill = ~survey_name, key := ~id) %>%
      add_tooltip( survey_label_cr, "hover") %>%
      add_axis("x", title = "Slide % Positive") %>%
      add_axis("y", title = "RDT % Positive") %>%
      scale_numeric("x", domain = c(0, 85)) %>%
      scale_numeric("y", domain = c(0, 85)) %>%
      layer_paths( data = reference_line, x = ~x, y= ~y)
    
# Regression    
  table.cr = reactive({
      linear.cr =  lm( rdt ~ slide, data = d_cr() ) 
      t = as.data.frame(cbind(round(linear.cr$coefficients, 2),
                                              confint(linear.cr))
                                        )
      names(t) = c("Estimate", "LL", "UL")
      rownames(t) = c("Intercept", "Slope")
      return(t)
  })

  renderTable( table.cr() )
  
    
```

## Interactive Country Chart 

```{r , echo=FALSE, message= FALSE, eval=FALSE}
# By REGION
    
renderText( "For all countries there is some disagreement between the population rates of RDT and microcopy.  For some, there is substantial deviation. See: Cote d'Ivoire and Tanzania as examples")

inputPanel(

  selectInput("Country", label = "Country",
              choices = country, selected = "")
)
 
    survey_label <- function(x) {
      dd = d_r()
      if(is.null(x)) return(NULL)
      # paste(nrow(dd))
      paste( dd[x$id, "country"], 
             "region", dd[x$id, "region"], 
             dd[x$id, "survey_name"], 
             "n=", dd[x$id, "n"])
    }

  
  d_r =  reactive({
    compare %>%
    # d[ country_selected, ] %>% 
    filter( (country %in% 
               ifelse( length(eval(input$Country))==0, 
                                         "Angola",
                                         eval(input$Country))) 
            ) %>%
    # filter( (country %in% country_selected) ) %>%
    group_by(country, region, survey_name) %>%
    summarise(
      n = sum(n),
      n_rdt = sum(n_rdt),
      n_slide = sum(n_slide),
      rdt = 100 * sum(rdt_pos) / n_rdt ,
      slide = 100* sum(slide_pos) / n_slide 
              ) %>%
    as.data.frame() %>%
    mutate( id = row_number()
            ) 
  })    

    d_r() %>% 
      group_by( survey_name) %>%
      ggvis(~slide, ~rdt ) %>%
      layer_model_predictions( model = "lm", se = TRUE
                               ) %>%
      ungroup() %>%
      layer_points( fill = ~survey_name, key := ~id) %>%
      add_tooltip( survey_label, "hover") %>%
      add_axis("x", title = "Slide % Positive") %>%
      add_axis("y", title = "RDT % Positive") %>%
      scale_numeric("x", domain = c(0, 85)) %>%
      scale_numeric("y", domain = c(0, 85)) %>%
      layer_paths( data = reference_line, x = ~x, y= ~y)

# Regression    
  table.r = reactive({
      linear.r =  lm( rdt ~ slide, data = d_r() ) 
      # t = as.data.frame(round(linear.r$coefficients, 2))
      t = as.data.frame(cbind(round(linear.r$coefficients, 2),
                                              confint(linear.r))
                                        )
      names(t) = c("Estimate", "LL", "UL")
      rownames(t) = c("Intercept", "Slope")
      return(t)
  })

  renderTable( table.r() )
  
      
```

# A Comparison of Test Standard Distribution 

## Cluster Standard Distribution within Region

Within each Region, the percent RDT and percent slide positive are calculated for each survey cluster.  Comparing the variance of these values may indicate the relative level of accuracy and repeatability of each method.


```{r, echo=FALSE, variance,  message= FALSE, eval=TRUE}
# 
 
  d_psu =  reactive({
    compare %>%
    filter( (country %in% 
               ifelse( length(eval(input$Country))==0, 
                                         "Angola",
                                         eval(input$Country))) 
            ) %>%
    group_by(country, survey_name, region, psu ) %>%
    summarise(
      n = sum(n),
      n_rdt = sum(n_rdt),
      n_slide = sum(n_slide),
      rdt = 100 * sum(rdt_pos) / n_rdt ,
      slide = 100* sum(slide_pos) / n_slide 
              ) %>%
    as.data.frame() %>%
   gather(test, perc_pos, rdt, slide) %>%
    mutate( id = row_number(),
      survey_name = paste(country, survey_name)) %>%
   arrange(country, survey_name, region, psu, test)
    
  })



  # ggvis 
#     d_psu %>% 
#       ggvis(~factor(region), ~test) %>%
#       ungroup() %>%
#       layer_points( fill = ~test, key := ~id) %>%
#       add_tooltip( survey_label, "hover") %>%
#       add_axis("x", title = "Region") %>%
#       add_axis("y", title = "% Positive") %>%
#       # scale_numeric("x", domain = c(0, 85)) %>%
#       scale_numeric("y", domain = c(0, 85)) 
  
  
   
renderPlot({

     a = d_psu() %>% 
       mutate(region = factor(region)) %>%
      ggplot( aes(region, perc_pos)) +
      geom_boxplot(aes(fill = test)) +
       facet_grid(survey_name ~ .)
    
     b = d_psu() %>% 
       mutate(region = factor(region) )%>%
       group_by(survey_name, region, test) %>%
       summarise(
         std = sd(perc_pos)
       ) %>%
      ggplot( aes(region, std)) +
      geom_point(aes(color = test), size = 4, 
                 position = position_jitter(w = 0.1, h = 0.1)) +
      facet_grid(survey_name ~ .)

  grid.arrange(a,b, ncol=1)    
})   

# renderTable(d_psu())
```

## Pooled Standard Distribution

```{r, echo=FALSE, message = FALSE, warning = FALSE, pooled_variance}

  d_psus =  
    compare %>%
    group_by(country, survey_name, region, psu ) %>%
    summarise(
      n = sum(n),
      n_rdt = sum(n_rdt),
      n_slide = sum(n_slide),
      rdt = 100 * sum(rdt_pos) / n_rdt ,
      slide = 100* sum(slide_pos) / n_slide 
              ) %>%
    as.data.frame() %>%
   gather(test, perc_pos, rdt, slide) %>%
    mutate( 
      id = row_number()
            ) %>%
   arrange(country, survey_name, region, psu, test)
    
# weighted std across regions 
  pooled_psu= 
    d_psus %>% 
       mutate(region = factor(region) ,
              test = factor(test)
              ) %>%
       group_by(country, survey_name, region, test) %>%
       summarise(
         n = sum(n),
         std = sd(perc_pos)
       ) %>%
      group_by(country, survey_name, test) %>%
    summarise(
      n = n(),
      pooled_sd = sqrt( sum(std^2 * (n - 1)) / (sum(n - 1)) )
    ) %>%
    mutate(
      id = paste(country, survey_name)
    )
    
# nrow(pooled_psu)
  
  ggplot(pooled_psu, aes(x = test, y= pooled_sd, color = test) ) +
           geom_point() + 
    geom_line( aes(group = id))
  
  pooled_psu %>% 
   spread(test, pooled_sd) %>%
  ggplot(aes(x = slide, y= rdt) ) + 
    geom_point() + 
    geom_smooth(method = "lm") +
    geom_abline(slope = 1 , intercept = 0)
    
# weighted std across all surveys 
  pooled_survey_sd= 
    pooled_psu %>% 
    group_by(test) %>%
    summarise(
      pooled_sd = sqrt( sum(pooled_sd^2 * (n - 1), na.rm = TRUE) / (sum(n - 1, na.rm = TRUE)) )
    ) 
    
  renderTable( pooled_survey_sd )
```




